PCRB

### 2013-02-21 doc/PCRB.tex
Fix minus sign typo
Fixed Bergman iteration
Updated CRLB calcualation to compute *posterior* bound not just a bound
of prior
Added discussion of sigmoid width
Also removed latex_includes.tex and fixed up PCRB document

\input{latex_includes.tex}

@@

The model takes the form
\begin{subequations}
\begin{align}
	\vec x_k &= \vec f(\vec x_{k-1}) + \vec v_k \\
	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
\end{align}
\label{eqn:model_general}
\end{subequations}
where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively. For this model the transition density is defined as
\begin{align}
	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
\end{align}
and the likelihood is
\begin{align}
	p(\vec y_k|\vec x_k) &= \mathcal N(\mat H\vec x_k,\mat R) \\
	&= (2\pi)^{-n/2} |\mat R|^{-1/2} \exp\left(-\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k) \right)
\end{align}

@@

The PCRB is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
\begin{align}
	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
\end{align}
where the expectation is taken over $\vec y$ and $\vec x$. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
\begin{align}
	\mat P &\ge \mat J^{-1}
\end{align}

@@

When noise is only added to a subset of the states each transition, $\mat Q$ is singular and $\mat Q^{-1}$ does not exist. Another method is required.

@@

An alternative formulation is to recursively compute the bound at each step. That is, compute $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$. The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g. multiplicative noise), denoted by a general transition function $\tilde{\vec f}(\vec x_k,\vec v_k)$. The recursion is
\begin{align}
	\mat P_k &= \bar{\mat F}_{k-1}\left(\mat P_{k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}\bar{\mat F}_{k-1} + \bar{\mat G}_{k-1}\mat Q_k \bar{\mat G}_{k-1}^{-1}
\end{align}
where
\begin{align}
	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
\end{align}
and $\mat P_0^{-1}$ is the inverse of the prior covariance. The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$. This simplifies the derivatives to
\begin{align}
	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
\end{align}
and the corresponding matrices to	
\begin{align}
	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
	\bar{\mat R}_k^{-1} &= -\mat H^T\mat R^{-1} \mat H \\
	\bar{\mat G}_k &= \mat I_{d\times d} 
\end{align}
Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.

@@

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{subfig}
\usepackage{float} % For [H] placement

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}

\usepackage{booktabs}

\usepackage[bookmarks=true,ocgcolorlinks=true,plainpages=false, breaklinks=true, bookmarksopen=true, bookmarksnumbered=true]{hyperref}
%\hypersetup{bookmarks=false}  %hide the bookmarks bar
%\hypersetup{bookmarksopen=false}  % expand tree of bookmarks or just show first level
\hypersetup{linkcolor=blue, citecolor=magenta,urlcolor=blue} % electronic
\hypersetup{colorlinks=true}

\usepackage[T1]{fontenc}
\usepackage{palatino}

\allowdisplaybreaks

\renewcommand{\vec}[1]{\ensuremath{{\boldsymbol #1}}}
\newcommand{\mat}[1]{\ensuremath{\boldsymbol{#1}}}

@@

The model takes the form
\begin{subequations}
	\label{eqn:model_general}
\begin{align}
	\vec x_k &= \vec f_k(\vec x_{k-1}) + \vec v_k \\
	\vec y_k &= \mat H\vec x_{k-1} + \vec w_k
\end{align}
\end{subequations}
where $\vec v_k$ and $\vec w_k$ are white, uncorrelated, zero mean Gaussian noise vectors with covariances of $\mat Q$ and $\mat R$, respectively. For this model the transition density is defined as
\begin{align}
	p(\vec x_k|\vec x_{k-1}) &= \mathcal N(\vec f(\vec x_{k-1})),\mat Q) \\
	&= (2\pi)^{-d/2} |\mat Q|^{-1/2} \exp\left(-\frac{1}{2}(\vec x_k - \vec f(\vec x_{k-1}))^T \mat Q^{-1} (\vec x_k - \vec f(\vec x_{k-1})) \right)
\end{align}
and the likelihood is
\begin{align}
	p(\vec y_k|\vec x_k) &= \mathcal N(\mat H\vec x_k,\mat R) \\
	&= (2\pi)^{-n/2} |\mat R|^{-1/2} \exp\left(-\frac{1}{2}(\vec y_k - \mat H\vec x_k)^T \mat R^{-1} (\vec y_k - \mat H\vec x_k) \right)
\end{align}

@@

The posterior Cram\'er-Rao bound (PCRB) is computed from the Fisher information matrix computed on the joint density $p(\vec y,\vec x)$,
\begin{align}
	\mat J &= \mathbb E\left[ \nabla_{\vec x}\nabla_{\vec x}^T \log p(\vec y,\vec x) \right]
\end{align}
where the expectation is taken over $\vec y$ and $\vec x$. The PCRB bounds the covariance of an unbiased estimator, $\mat P$,
\begin{align}
	\mat P &\ge \mat J^{-1}
\end{align}

@@

This method is not applicable when noise is only added to a subset of the states each transition. In this case, $\mat Q$ is singular and $\mat Q^{-1}$ does not exist. Another method is required.

@@

An alternative formulation is to recursively compute the bound instead of the matrix, i.e., $\mat P_k = \mat J_k^{-1}$ directly rather than $\mat J_k$. The formulas suggested in \cite{Bergman2001} are applicable to wider class of models (e.g. multiplicative noise), denoted by a general transition function $\tilde{\vec f}(\vec x_k,\vec v_k)$. The recursion is
\begin{subequations}
\begin{align}
	\mat P_{k|k-1} &= \bar{\mat F}_k\mat P_{k-1|k-1}\bar{\mat F}_k + \bar{\mat G}_k\mat Q_k \bar{\mat G}_k^{-1} \\
		\mat P_{k|k} &= \left(\mat P_{k|k-1}^{-1} + \bar{\mat R}_{k-1}^{-1}\right)^{-1}
\end{align}
\end{subequations}
where
\begin{align}
	\bar{\mat F}_k &= \mathbb E \left[ \nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\bar{\mat R}_k^{-1} &= -\mathbb E\left[ \nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) \right] \\
	\bar{\mat G}_k &= \mathbb E \left[ \nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k)\right] \\
	\mat Q_k^{-1} &= -\mathbb E \left[ \nabla_{\vec v_k}\nabla_{\vec v_k}^T \log p(\vec v_k) \right]
\end{align}
and $\mat P_0^{-1}$ is the inverse of the prior covariance. The model in \eqref{eqn:model_general} is represented by $\tilde{\vec f}(\vec x_k,\vec v_k) = \vec f(\vec x_k) + \vec v_k$. This simplifies the derivatives to
\begin{align}
	\nabla_{\vec x_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat F_k(\vec x_k) \\
	\nabla_{\vec v_k}^T \tilde{\vec f_k}(\vec x_k,\vec v_k) &= \mat I_{d\times d} \\
	\nabla_{\vec x_k}\nabla_{\vec x_k}^T \log p(\vec y_k|\vec x_k) &= -\mat H^T\mat R^{-1} \mat H 
\end{align}
and the corresponding matrices to	
\begin{align}
	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \label{eqn:pcrb_term_F}\\
	\bar{\mat R}_k^{-1} &= \mat H^T\mat R^{-1} \mat H \\
	\bar{\mat G}_k &= \mat I_{d\times d} 
\end{align}
Notice that $\mat Q^{-1}$ is not directly required in the recursion, so this method is applicable to neural field models where process noise is only added to a single state.

@@

\subsection{Analytical evaluation of expectations}

For the JR model the expectation in \eqref{eqn:pcrb_term_F} above is
\begin{align}
	\bar{\mat F}_k &= \mathbb E\left[ \mat F_k(\vec x_k)\right] \\
	&= \mat F_k^{\rm (lin)} + \mathbb E\left[ \mat G_k(\vec x_k)\right] 
\end{align}
where $\mat G_k(\vec x_k)$ contains the nonlinear component of the Jacobian. These nonlinear terms are derivatives of the error function and have the form,
\begin{align}
	\exp(-(x-v_0)^2/\zeta^2)
\end{align}
The expectation of such terms can only be evaluated when we know the prior mean and covariance at all times.
